\documentclass[journal]{IEEEtran}

% Math packages
\usepackage{amsmath,amsfonts,amssymb,amsthm,bm,mathrsfs}

% Algorithm packages
\usepackage{algorithmic,algorithm}

% Array and table packages
\usepackage{array,multirow,booktabs,makecell}

% Graphics packages
\usepackage{graphicx,epsfig}

% Float and sizing packages
\usepackage{stfloats,float,calc}

% Text and formatting packages
\usepackage{textcomp,verbatim,url,cite,color,xcolor,cancel}
% Changes package for revision highlights
\let\comment\undefined
\usepackage[draft]{changes}
\definechangesauthor[name={cuiys}, color=red]{cuiys}
% 使用示例：\added[id=cuiys]{新增内容}，\deleted[id=cuiys]{删除内容}，\replaced[id=cuiys]{新}{旧}

% Layout packages
\usepackage{cuted}

\hyphenation{op-tical net-works semi-conduc-tor IEEE-Xplore}

% Environment customization
\makeatletter
\providecommand{\tabularnewline}{\\}
\floatstyle{ruled}
\newfloat{algorithm}{tbp}{loa}
\providecommand{\algorithmname}{Algorithm}
\floatname{algorithm}{\protect\algorithmname}
\makeatother

% Theorem environments
\theoremstyle{plain}
\newtheorem{theorem}{Theorem}
\newtheorem{assumption}{Assumption}
\newtheorem{lemma}{Lemma}
\newtheorem{lem}{\protect\lemmaname}
\newtheorem{thm}{\protect\theoremname}
\newtheorem{pro}{\protect\proposition}
\newtheorem{corollary}{Corollary}

\theoremstyle{definition}
\newtheorem{defn}{\protect\definitionname}
\newtheorem{myDef}{Definition}

% Custom command definitions
\providecommand{\definitionname}{Definition}
\providecommand{\lemmaname}{Lemma}
\providecommand{\theoremname}{Theorem}
\providecommand{\proposition}{Proposition}
\renewcommand{\algorithmicrequire}{\textbf{Input:}}
\renewcommand{\algorithmicensure}{\textbf{Output:}}

% Display settings
\allowdisplaybreaks[4]

\begin{document}

\title{Satellite TT\&C System Federated Fault Diagnosis:\\Architecture Design and Convergence Analysis}

\author{\IEEEauthorblockN{Yongsheng Cui\IEEEauthorrefmark{1}, Taiyi Chen\IEEEauthorrefmark{1}, Haoran Xie\IEEEauthorrefmark{2}, and Yafeng Zhan\IEEEauthorrefmark{1,3}} \\
\IEEEauthorblockA{\IEEEauthorrefmark{1}Department of Electronic Engineering, Tsinghua University, Beijing, China\\
\IEEEauthorrefmark{2}Electric Power Research Institute, China Southern Power Grid, Guangzhou, China\\
\IEEEauthorrefmark{3}Beijing National Research Center for Information Science and Technology, Tsinghua University, Beijing, China}
\thanks{This work was supported by the National Natural Science Foundation of China under Grant 62131012.}
\thanks{Y. Zhan is the corresponding author (e-mail: zhanyf@mail.tsinghua.edu.cn).}}

\maketitle

\begin{abstract}
Satellites operate in harsh space environments for extended periods, 
making fault diagnosis a critical strategy for extending their lifespan. 
Federated learning based fault diagnosis using telemetry data addresses critical challenges in satellite system reliability, 
providing solutions for data privacy across different telemetry, tracking and command (TT\&C) centers 
and improving fault diagnosis performance when individual TT\&C centers have limited training data. 
However, challenges arise in the deployment and application of federated learning methods
due to \textit{``Communication-limited''} and \textit{``Data Heterogeneity''} at some TT\&C centers, 
causing performance degradation and slow convergence.
To solve these problems, this paper proposes a streaming federated learning framework for satellite fault diagnosis. 
The framework incorporates an Incremental Learning Algorithm based on Dual-Domain Knowledge Distillation for efficient model updates, 
and a Federated Learning Algorithm based on Partial Forgetting Compensation for handling data heterogeneity. 
Under this framework, TT\&C centers with communication constraints only need to participate in a single round of federated learning 
for the system to obtain a well-performing and stable model. 
Theoretical convergence analysis is provided to characterize the algorithm performance. 
Extensive simulations show that the proposed algorithm outperforms existing fault diagnosis approaches for satellite systems.


\end{abstract}

\begin{IEEEkeywords}
  Communication constraints, convergence analysis, data heterogeneity, federated learning, knowledge distillation, satellite fault diagnosis.
\end{IEEEkeywords}

\input{chapter/Introduction.tex}
\input{chapter/Related Work.tex}
\input{chapter/Proposed Streaming Federated Learning Framework.tex}
% \input{chapter/Convergence Analysis.tex}
\input{chapter/Convergence_Analysis_Draft.tex}
\input{chapter/Simulation Results.tex}
\input{chapter/Conclusion.tex}
% \input{chapter/Appendix.tex}
\input{chapter/Appendix_Draft.tex}
\end{document}


